### 2023-02-08 template.Rmd
autogit

---
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Statutory Declaration

I herewith declare that I have composed the present thesis myself and without use of any other than the
cited sources and aids. Sentences or parts of sentences quoted literally are marked as such; other references
with regard to the statement and scope are indicated by full details of the publications concerned. The thesis
in the same or similar form has not been submitted to any examination body and has not been published.
This thesis was not yet, even in part, used in another examination or as a course performance

Place, Date:  
Signature:

## Ehrenwörtliche Erklärung

Ich versichere hiermit, dass ich die vorliegende Arbeit selbständig und ohne Benutzung anderer als der
angegebenen Quellen und Hilfsmittel verfasst habe. Wörtlich übernommene Sätze oder Satzteile sind als
Zitat belegt, andere Anlehnungen, hinsichtlich Aussage und Umfang, unter Quellenangabe kenntlich
gemacht. Die Arbeit hat in gleicher oder ähnlicher Form noch keiner Prüfungsbehörde vorgelegen und ist
nicht veröffentlicht. Sie wurde nicht, auch nicht auszugsweise, für eine andere Prüfungs- oder
Studienleistung verwendet.

Ort, Datum:  
Unterschrift: