### 2023-02-15 statutory_declaration/statutory_declaration.Rmd
autogit

---
output: pdf_document
pagestyle: empty
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Statutory Declaration

I hereby assure that I have independently and without using sources or resources other than those indicated, written this paper.
Directly quoted sentences or sentence parts are marked as citations, while other sources of inspiration, in terms of content and scope, are identified with references.
The work has not been presented to any examination board in the same or a similar form and has not been published.
It has not been used, even partially, for any other examination or academic achievement.
I am aware of the regulations for ensuring good scientific practice in research and teaching, as well as the procedures for dealing with scientific misconduct.

Place, Date:

------------------------------------------------------------------------

Signature:

------------------------------------------------------------------------

## Ehrenwörtliche Erklärung

Ich versichere hiermit, dass ich die vorliegende Arbeit selbständig und ohne Benutzung anderer als der angegebenen Quellen und Hilfsmittel verfasst habe.
Wörtlich übernommene Sätze oder Satzteile sind als Zitat belegt, andere Anlehnungen, hinsichtlich Aussage und Umfang, unter Quellenangabe kenntlich gemacht.
Die Arbeit hat in gleicher oder ähnlicher Form noch keiner Prüfungsbehörde vorgelegen und ist nicht veröffentlicht.
Sie wurde nicht, auch nicht auszugsweise, für eine andere Prüfungs- oder Studienleistung verwendet.
Von der Ordnung zur Sicherung guter wissenschaftlicher Praxis in Forschung und Lehre und zum Verfahren zum Umgang mit wissenschaftlichem Fehlverhalten wurde Kenntnis genommen wurde.

Ort, Datum:

------------------------------------------------------------------------

Unterschrift:

------------------------------------------------------------------------